Optimal synaptic strategies for different timescales of memory

### 2018-03-16 Paper/SynapsePaper.tex
cleaning

%% ADDITIONAL OPTIONAL STYLE FILES
%\usepackage[hypcap]{caption}%links to figures
\usepackage{graphicx}
%\usepackage{graphics} for finding documentation only
\pdfsuppresswarningpagegroup=1
\usepackage{epstopdf}
\epstopdfsetup{update,suffix=-generated}
%
\IfFileExists{"C:/Progra\string~1/Inkscape/inkscape.exe"}{\newcommand{\inkscapedir}{C:/Progra\string~1/Inkscape/}}{}
\IfFileExists{"/Applications/Inkscape.app/Contents/Resources/bin/inkscape"}{\newcommand{\inkscapedir}{/Applications/Inkscape.app/Contents/Resources/bin/}}{}
\providecommand{\inkscapedir}{}
%
\epstopdfDeclareGraphicsRule{.svg}{pdf}{.pdf}{"\inkscapedir inkscape" -f #1 -C -A \OutputFile}
%
\usepackage{adjustbox}
\newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
%
%\usepackage{paralist}
%\newenvironment{myenuma}{\begin{inparaenum}[(a)]}{\end{inparaenum}}
%\newenvironment{myenumA}{\begin{inparaenum}[\bfseries A.]}{\end{inparaenum}}
%
%\usepackage[inline]{enumitem}
\newlist{myenuma}{enumerate*}{4}
\setlist[myenuma]{label=(\alph*),ref=\alph*}
\newcommand{\itemp}[1]{\item\aligntop{#1}}

\graphicspath{{"Figs/"}}
\usepackage{amssymb,amsfonts}
%\usepackage[centertags]{amsmath}
%\usepackage{hyperref}
\usepackage[capitalise,nameinlink]{cleveref}
\creflabelformat{equation}{[#2#1#3]}
\crefrangelabelformat{equation}{[#3#1#4--#5#2#6]}
%\crefformat{equation}{#2Eq.~[#1]#3}
%\crefrangeformat{equation}{Eqs.~[#3#1#4--#5#2#6]}
%\crefmultiformat{equation}{Eqs.~[#2#1#3}{,2#1#3]}{,#2#1#3}{,#2#1#3]}
%\crefrangemultiformat{equation}{Eqs.~[#3#1#4--#5#2#6}{,#3#1#4--#5#2#6]}{,#3#1#4--#5#2#6}{,#3#1#4--#5#2#6]}

%% OPTIONAL MACRO DEFINITIONS
\newcommand{\ra}{\rightarrow}
\newcommand{\lr}{\leftrightarrow}
\newcommand{\cdt}{\!\cdot\!}
\newcommand{\vp}{\vspace{0.5cm}}
\newcommand{\hp}{\hspace{0.5cm}}
%
%e.g., i.e. with normal spaces
\newcommand{\eg}{e.g.\ }
\newcommand{\ie}{i.e.\ }
\newcommand{\cf}{cf.\ }
\newcommand{\etc}{etc.\ }
\newcommand{\wrt}{wrt.\ }
\newcommand{\Eg}{E.g.\ }
\newcommand{\Ie}{I.e.\ }
\newcommand{\Cf}{Cf.\ }
\newcommand{\Etc}{Etc.\ }
%
\usepackage{delim}
% brackets etc.
\delimdef\prn#1{\dleft( #1 \dright)}
\delimdef\brc#1{\dleft\{ #1 \dright\}}
\delimdef\brk#1{\dleft[ #1 \dright]}
\delimdef\abs#1{\dleft\lvert #1 \dright\rvert}
\delimdef\nrm#1{\dleft\lVert #1 \dright\rVert}
\delimdef\av#1{\dleft\langle #1 \dright\rangle}
\delimdef\avc#1#2{\dleft\langle #1 \dmiddle\vert #2 \dright\rangle}
%\newenvironment{cases}{\dleft\{\begin{aligned}}{\end{aligned}\dright.}
%
% Sets
\delimdef\set#1#2{\dleft\{ #1 \dmiddle\vert #2 \dright\}}
\delimdef\cond#1#2{\!\dleft( #1 \dmiddle\vert #2 \dright)}
\delimdef\condb#1#2{\!\dleft[ #1 \dmiddle\vert #2 \dright]}
%
% Derivatives, etc. First argument is optional.
\newcommand{\diff}[3][]{\frac{\mathrm{d}^{#1} #2}{\mathrm{d}{#3}^{#1}}}
\newcommand{\pdiff}[3][]{\frac{\partial^{#1} #2}{\partial {#3}^{#1}}}
\newcommand{\pdiffc}[3][]{\left(\frac{\partial #2}{\partial {#3}}\right)_{\!\!#1}}
\newcommand{\intd}[2][]{\int #1\!\!\dr #2\,}
%
% Un-italicised letters
\newcommand{\dr}{\mathrm{d}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\ir}{\mathrm{i}}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator{\Det}{Det}
\DeclareMathOperator{\cov}{Cov}
\DeclareMathOperator{\var}{Var}
\DeclareMathOperator{\bias}{Bias}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\sech}{sech}
\DeclareMathOperator{\diag}{diag}
%
% The default \Im and \Re look crap
\renewcommand{\Im}{\operatorname{\mathfrak{Im}}}
\renewcommand{\Re}{\operatorname{\mathfrak{Re}}}
%
% logic
\newcommand{\means}{\Longleftrightarrow}
\newcommand{\requires}{\Longleftarrow}
%
%\input{sl_symbols.tex}
%
%additional symbols:
\newcommand{\half}{\frac{1}{2}}
\newcommand{\CO}{\mathcal{O}}
\newcommand{\CL}{\mathcal{L}}
\newcommand{\dt}{\dr t}
%
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
\newcommand{\wv}{\vec{w}}
\newcommand{\wvi}{\vec{w}_\text{ideal}}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\dg}{^\mathrm{dg}}
\newcommand{\trans}{^\mathrm{T}}
\newcommand{\I}{\mathbf{I}}
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MM}{\mathbf{Q}}
%prob distributions
\newcommand{\prob}{\mathbf{p}}
\newcommand{\eq}{\prob^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fund}{\mathbf{Z}}
%other symbols for matrices
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Lb}{\boldsymbol{\Lambda}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\enc}{\mathbf{q}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\F}{\boldsymbol{\Phi}}
%superscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
%sets
\newcommand{\CS}{\mathcal{S}}
\newcommand{\CA}{\mathcal{A}}
\newcommand{\CB}{\mathcal{B}}
\newcommand{\comp}{^\mathrm{c}}
%eigenmodes
\newcommand{\uv}{\mathbf{u}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\CI}{\mathcal{I}}

% ----------------------------------------------------------------
\input{sl_theorems_preamble.tex}
% ----------------------------------------------------------------
\providecommand\hyperref[2][1]{#2}
\providecommand\texorpdfstring[2]{#1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Optimal synaptic strategies for different timescales of memory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1,2]{Subhaneil Lahiri}
\author[b,1]{Surya Ganguli}

@@

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{article}

@@

\begin{figure}[tbp]
 \begin{center}
 \begin{myenuma}
  \protect\item\aligntop{\includegraphics[width=0.25\linewidth]{cascade.svg}}\label{fig:cascade}
  \item\aligntop{\includegraphics[width=0.25\linewidth]{multistate.svg}}\label{fig:multistate}
  \item\aligntop{\includegraphics[width=0.3\linewidth]{cascms.eps}}\label{fig:cascms}
 \end{myenuma}
 \end{center}
  \caption{Models of complex synapses.
  (\ref{fig:cascade}) The cascade model of \cite{Fusi2005cascade}, showing transitions between states of high/low synaptic weight (red/blue circles) due to potentiation/depression (solid red/dashed blue arrows).
  (\ref{fig:multistate}) The serial model of \cite{Leibold2008serial}.
  (\ref{fig:cascms}) The memory curves of these two models, showing the decay of the signal-to-noise ratio (to be defined \hyperref[sec:setup]{below}) as subsequent memories are stored.
  \label{fig:models}}
\end{figure}

@@

This two state model has the equilibrium distribution $\eq=(f\dep,f\pot)$ and its flux is given by $\F_{-+} = rf\pot f\dep$.
This is maximized when $f\pot=f\dep=\half$, leading to the upper bound:
%
\begin{equation}\label{eq:max_init}
  \snr(0) \leq \sqrt{N}.
\end{equation}
%
We note that while this model has high initial SNR, it also has very fast memory decay -- with a timescale $\tau\sim\frac{1}{r}$.
As the synapse is very plastic, the initial memory is encoded very easily, but the subsequent memories also overwrite it rapidly.
This is one example of the tradeoff between optimizing memory at early versus late times.
\begin{figure}[tbp]
 \begin{center}
 \begin{myenuma}
  \item\hp\aligntop{\includegraphics[width=0.24\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
  \item\aligntop{\includegraphics[width=0.24\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
  \item\hp\aligntop{\includegraphics[width=0.1\linewidth]{binary_det.svg}}\label{fig:binary_det}
  \end{myenuma}
 \end{center}
  \caption{Synaptic models that maximize initial SNR.
  (\ref{fig:max_init_pot}) For potentiation, all transitions starting from a weak state lead to a strong state, and the probabilities for all transitions leaving a given weak state sum to 1.
  (\ref{fig:max_init_pot}) Depression is similar to potentiation, but with strong and weak interchanged.
  (\ref{fig:binary_det}) The equivalent two state model, with transition probabilities under potentiation and depression equal to one.
  \label{fig:max_init}}
\end{figure}

@@

In the context of complex synapses, we can define the partial sums
%
\begin{equation}\label{eq:kemenypm}
   \eta^+_i = \sum_{j\in+} \fptb_{ij} \eq_j,
   \qquad
   \eta^-_i = \sum_{j\in-} \fptb_{ij} \eq_j.
\end{equation}
%
These can be thought of as the average time it takes to reach the strong/weak states respectively.
Using these definitions, we can then impose an order on the states by arranging them in order of decreasing $\smash{\eta^+_i}$ or increasing $\smash{\eta^-_i}$.
Because $\smash{\eta^+_i+\eta^-_i=\eta}$ is independent of $i$, the two orderings are the same.
In this order, which depends sensitively on the structure of $\M\potdep$, states later (to the right in figures below) can be considered to be more potentiated than states earlier (to the left in figures below), despite the fact that they have the same synaptic efficacy.
In essence, in this order, a state is considered to be more potentiated if the average time it takes to reach all the strong efficacy states is shorter.
We will see that synaptic models that optimize various measures of memory have an exceedingly simple structure when, and only when, their states are arranged in this order.%
\footnote{Note that we do not need to worry about the order of the $\eta^\pm_i$ changing during the optimization: necessary conditions for a maximum only require that there is no infinitesimal perturbation that increases the area. Therefore we need only consider an infinitesimal neighborhood of the model, in which the order will not change.}

@@

\begin{figure}[tbp]
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=0.14\linewidth]{triangular_right.svg}}\label{fig:tri_right}\hp
  \item\aligntop{\includegraphics[width=0.14\linewidth]{triangular_left.svg}}\label{fig:tri_left}\hp
  \item\aligntop{\includegraphics[width=0.28\linewidth]{shortcut.svg}}\label{fig:shortcut}
 \end{myenuma}
 \end{center}
  \caption{Perturbations that increase the area.
  (\ref{fig:tri_right}) Perturbations that increase elements of $\M\pot$ above the diagonal and decrease the corresponding elements of $\M\dep$. It can no longer be used when $\M\dep$ is lower triangular, \ie depression must move synapses to ``more depressed'' states.
  (\ref{fig:tri_left}) Perturbations that decrease elements of $\M\pot$ below the diagonal and increase the corresponding elements of $\M\dep$. It can no longer be used when $\M\pot$ is upper triangular, \ie potentiation must move synapses to ``more potentiated'' states.
  (\ref{fig:shortcut}) Perturbation that decreases ``shortcut'' transitions and increases the bypassed ``direct'' transitions. It can no longer be used when there are only nearest-neighbor ``direct'' transitions.
  \label{fig:perts}}
\end{figure}

@@

\begin{figure}
 \begin{center}
% \parbox[t]{0.43\linewidth}{%
 \begin{myenuma}
   \item\aligntop{\includegraphics[width=0.85\linewidth]{env.svg}}\label{fig:envplot}
% \end{myenuma}}\hp
% \parbox[t]{0.44\linewidth}{%
% \begin{myenuma}
%  \stepcounter{myenumai}
%% \addtocounter{enumi}{1}
   \item\aligntop{\includegraphics[width=0.8\linewidth]{diffjump.svg}}\label{fig:diffjump}%\\[0.1\linewidth]
   \item\aligntop{\includegraphics[width=0.8\linewidth]{multistate_sticky.svg}}\label{fig:sticky}
 \end{myenuma}%}
 \end{center}
  \caption{The memory curve envelope for $N=100$, $M=12$.
  (\protect\protect\ref{fig:envplot}) An upper bound on the SNR at any time is shown in green.
  The red dashed curve shows the result of numerical optimization of synaptic models with random initialization.
  The solid red curve shows the highest SNR we have found with hand designed models.
  At early times these models are of the form shown in (\ref{fig:diffjump}) with different numbers of states, and all transition probabilities equal to $1$.
  At late times they are of the form shown in (\ref{fig:sticky}) with different values of $\varepsilon$.
  The model shown in (\ref{fig:sticky}) also saturates the area bound \cref{eq:max_area} in the limit $\varepsilon\to0$.
  \label{fig:env}}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\bibliographystyle{pnas-new}
\bibliography{maths,neuro,markov}

%\end{article}
\clearpage

\begin{strip}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\appendix
\begin{center}
  \textbf{\Large Supplementary material}
\end{center}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In the \hyperref[sec:areacoeffincr]{section below}, we will show that the $c_k$ are non-decreasing, if we put the states in order of decreasing $\eta^+_k$.
This implies that the last term of the final expression in \cref{eq:shortcutarea} is non-negative.
If it is non-zero (there would need to be a lot of degeneracy for it to be zero), this would imply that $\mu\potdep _{g,g\pm m}>0$, which in turn implies that $\W\potdep _{g,g\pm m}=0$.
This would tell us that the process with the maximal area has to have a multi-state topology.

@@

In the \hyperref[sec:shortcuts]{previous subsection} we defined perturbations corresponding to Figure 3(c) of the main text.
In order to show that those perturbations increase the area, we must now show that the $c_k$ are non-decreasing, if we put the states in order of decreasing $\eta^+_k$.

@@

%\bibliographystyle{utcaps_sl}
%\bibliography{maths,neuro,markov}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% ADDITIONAL OPTIONAL STYLE FILES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{graphicx}
%\usepackage{graphics} for finding documentation only
\pdfsuppresswarningpagegroup=1
\usepackage{epstopdf}
\epstopdfsetup{update,suffix=-generated}
%
\IfFileExists{"C:/Progra\string~1/Inkscape/inkscape.exe"}{\newcommand{\inkscapedir}{C:/Progra\string~1/Inkscape/}}{}
\IfFileExists{"/Applications/Inkscape.app/Contents/Resources/bin/inkscape"}{\newcommand{\inkscapedir}{/Applications/Inkscape.app/Contents/Resources/bin/}}{}
\providecommand{\inkscapedir}{}
%
\epstopdfDeclareGraphicsRule{.svg}{pdf}{.pdf}{"\inkscapedir inkscape" -f #1 -C -A \OutputFile}
%
\usepackage{adjustbox}
\newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
%
%\usepackage[inline]{enumitem}
\newlist{myenuma}{enumerate*}{4}
\setlist[myenuma]{label=(\alph*),ref=\alph*}
\newcommand{\panel}[1]{\item\aligntop{#1}}
\graphicspath{{"Figs/"}}

\usepackage{amssymb,amsfonts}
\usepackage{amsthm}
%\usepackage[centertags]{amsmath}
%\usepackage{hyperref}
\usepackage[capitalise,nameinlink]{cleveref}
\creflabelformat{equation}{[#2#1#3]}
\crefrangelabelformat{equation}{[#3#1#4--#5#2#6]}
%\crefformat{equation}{#2Eq.~[#1]#3}
%\crefrangeformat{equation}{Eqs.~[#3#1#4--#5#2#6]}
%\crefmultiformat{equation}{Eqs.~[#2#1#3}{,2#1#3]}{,#2#1#3}{,#2#1#3]}
%\crefrangemultiformat{equation}{Eqs.~[#3#1#4--#5#2#6}{,#3#1#4--#5#2#6]}{,#3#1#4--#5#2#6}{,#3#1#4--#5#2#6]}

% ----------------------------------------------------------------
\newtheoremstyle{sldefinition}%
  {3pt}%space above
  {3pt}%space below
  {}%body font
  {}%indent amount
  {\bfseries}%theorem head font
  {}%theorem head punctuation
  {\newline}%space after head
  {\thmname{#1}\thmnumber{ #2}:{\bfseries\thmnote{ #3}}}%head spec
\newtheoremstyle{slplain}%
  {3pt}%space above
  {3pt}%space below
  {}%body font
  {}%indent amount
  {\bfseries}%theorem head font
  {}%theorem head punctuation
  {\newline}%space after head
  {\thmname{#1}\thmnumber{ #2}{\bfseries\thmnote{ (#3)}}: }%head spec
%
\theoremstyle{slplain}
\newtheorem{thm}{Theorem}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
%
\theoremstyle{sldefinition}
\newtheorem{defn}{Definition}
%
\theoremstyle{remark}
\newtheorem*{notn}{Notation}
\newtheorem*{rem}{Remark}
% ----------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% OPTIONAL MACRO DEFINITIONS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\ra}{\rightarrow}
\newcommand{\lr}{\leftrightarrow}
\newcommand{\cdt}{\!\cdot\!}
\newcommand{\vp}{\vspace{0.5cm}}
\newcommand{\hp}{\hspace{0.5cm}}
%
%e.g., i.e. with normal spaces
\newcommand{\eg}{e.g.\ }
\newcommand{\ie}{i.e.\ }
\newcommand{\cf}{cf.\ }
\newcommand{\etc}{etc.\ }
\newcommand{\wrt}{wrt.\ }
\newcommand{\Eg}{E.g.\ }
\newcommand{\Ie}{I.e.\ }
\newcommand{\Cf}{Cf.\ }
\newcommand{\Etc}{Etc.\ }
%
\usepackage{delim}
% brackets etc.
\delimdef\prn#1{\dleft( #1 \dright)}
\delimdef\brc#1{\dleft\{ #1 \dright\}}
\delimdef\brk#1{\dleft[ #1 \dright]}
\delimdef\abs#1{\dleft\lvert #1 \dright\rvert}
\delimdef\nrm#1{\dleft\lVert #1 \dright\rVert}
\delimdef\av#1{\dleft\langle #1 \dright\rangle}
\delimdef\avc#1#2{\dleft\langle #1 \dmiddle\vert #2 \dright\rangle}
%\newenvironment{cases}{\dleft\{\begin{aligned}}{\end{aligned}\dright.}
%
% Sets
\delimdef\set#1#2{\dleft\{ #1 \dmiddle\vert #2 \dright\}}
\delimdef\cond#1#2{\!\dleft( #1 \dmiddle\vert #2 \dright)}
\delimdef\condb#1#2{\!\dleft[ #1 \dmiddle\vert #2 \dright]}
%
% Derivatives, etc. First argument is optional.
\newcommand{\diff}[3][]{\frac{\mathrm{d}^{#1} #2}{\mathrm{d}{#3}^{#1}}}
\newcommand{\pdiff}[3][]{\frac{\partial^{#1} #2}{\partial {#3}^{#1}}}
\newcommand{\pdiffc}[3][]{\left(\frac{\partial #2}{\partial {#3}}\right)_{\!\!#1}}
\newcommand{\intd}[2][]{\int #1\!\!\dr #2\,}
%
% Un-italicised letters
\newcommand{\dr}{\mathrm{d}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\ir}{\mathrm{i}}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator{\Det}{Det}
\DeclareMathOperator{\cov}{Cov}
\DeclareMathOperator{\var}{Var}
\DeclareMathOperator{\bias}{Bias}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\sech}{sech}
\DeclareMathOperator{\diag}{diag}
%
% The default \Im and \Re look crap
\renewcommand{\Im}{\operatorname{\mathfrak{Im}}}
\renewcommand{\Re}{\operatorname{\mathfrak{Re}}}
%
% logic
\newcommand{\means}{\Longleftrightarrow}
\newcommand{\requires}{\Longleftarrow}
%
%\input{sl_symbols.tex}
%
%additional symbols:
\newcommand{\half}{\frac{1}{2}}
\newcommand{\CO}{\mathcal{O}}
\newcommand{\CL}{\mathcal{L}}
\newcommand{\dt}{\dr t}
%
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
\newcommand{\wv}{\vec{w}}
\newcommand{\wvi}{\vec{w}_\text{ideal}}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\dg}{^\mathrm{dg}}
\newcommand{\trans}{^\mathrm{T}}
\newcommand{\I}{\mathbf{I}}
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MM}{\mathbf{Q}}
%prob distributions
\newcommand{\prob}{\mathbf{p}}
\newcommand{\eq}{\prob^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fund}{\mathbf{Z}}
%other symbols for matrices
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Lb}{\boldsymbol{\Lambda}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\enc}{\mathbf{q}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\F}{\boldsymbol{\Phi}}
%superscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
%sets
\newcommand{\CS}{\mathcal{S}}
\newcommand{\CA}{\mathcal{A}}
\newcommand{\CB}{\mathcal{B}}
\newcommand{\comp}{^\mathrm{c}}
%eigenmodes
\newcommand{\uv}{\mathbf{u}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\CI}{\mathcal{I}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Optimal synaptic strategies for different timescales of memory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1,2]{Subhaneil Lahiri}
\author[b,1]{Surya Ganguli}

@@

\setboolean{displaywatermark}{false}
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{article}

@@

\begin{figure}[tbp]
 \begin{center}
 \begin{myenuma}
  \panel{\includegraphics[width=0.25\linewidth]{cascade.svg}}\label{fig:cascade}
  \panel{\includegraphics[width=0.25\linewidth]{multistate.svg}}\label{fig:multistate}
  \panel{\includegraphics[width=0.3\linewidth]{cascms.eps}}\label{fig:cascms}
 \end{myenuma}
 \end{center}
  \caption{Models of complex synapses.
  (\ref{fig:cascade}) The cascade model of \cite{Fusi2005cascade}, showing transitions between states of high/low synaptic weight (red/blue circles) due to potentiation/depression (solid red/dashed blue arrows).
  (\ref{fig:multistate}) The serial model of \cite{Leibold2008serial}.
  (\ref{fig:cascms}) The memory curves of these two models, showing the decay of the signal-to-noise ratio (to be defined \hyperref[sec:setup]{below}) as subsequent memories are stored.
  \label{fig:models}}
\end{figure}

@@

This two state model has the equilibrium distribution $\eq=(f\dep,f\pot)$ and its flux is given by $\F_{-+} = rf\pot f\dep$.
This is maximized when $f\pot=f\dep=\half$, leading to the upper bound:
%
\begin{equation}\label{eq:max_init}
  \snr(0) \leq \sqrt{N}.
\end{equation}
%
We note that while this model has high initial SNR, it also has very fast memory decay -- with a timescale $\tau\sim\frac{1}{r}$.
As the synapse is very plastic, the initial memory is encoded very easily, but the subsequent memories also overwrite it rapidly.
This is one example of the tradeoff between optimizing memory at early versus late times.
\begin{figure}[tbp]
 \begin{center}
 \begin{myenuma}
  \panel{\hp\includegraphics[width=0.24\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
  \panel{\includegraphics[width=0.24\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
  \panel{\hp\includegraphics[width=0.1\linewidth]{binary_det.svg}}\label{fig:binary_det}
  \end{myenuma}
 \end{center}
  \caption{Synaptic models that maximize initial SNR.
  (\ref{fig:max_init_pot}) For potentiation, all transitions starting from a weak state lead to a strong state, and the probabilities for all transitions leaving a given weak state sum to 1.
  (\ref{fig:max_init_pot}) Depression is similar to potentiation, but with strong and weak interchanged.
  (\ref{fig:binary_det}) The equivalent two state model, with transition probabilities under potentiation and depression equal to one.
  \label{fig:max_init}}
\end{figure}

@@

In the context of complex synapses, we can define the partial sums
%
\begin{equation}\label{eq:kemenypm}
   \eta^+_i = \sum_{j\in+} \fptb_{ij} \eq_j,
   \qquad
   \eta^-_i = \sum_{j\in-} \fptb_{ij} \eq_j.
\end{equation}
%
These can be thought of as the average time it takes to reach the strong/weak states respectively.
Using these definitions, we can then impose an order on the states by arranging them in order of decreasing $\smash{\eta^+_i}$ or increasing $\smash{\eta^-_i}$.
Because $\smash{\eta^+_i+\eta^-_i=\eta}$ is independent of $i$, the two orderings are the same.
In this order, which depends sensitively on the structure of $\M\potdep$, states later (to the right in figures below) can be considered to be more potentiated than states earlier (to the left in figures below), despite the fact that they have the same synaptic efficacy.
In essence, in this order, a state is considered to be more potentiated if the average time it takes to reach all the strong efficacy states is shorter.
We will see that synaptic models that optimize various measures of memory have an exceedingly simple structure when, and only when, their states are arranged in this order.%
\footnote{Note that we do not need to worry about the order of the $\eta^\pm_i$ changing during the optimization: necessary conditions for a maximum only require that there is no infinitesimal perturbation that increases the area. Therefore we need only consider an infinitesimal neighborhood of the model, in which the order will not change.\label{ft:etaorder}}

@@

\begin{figure}[tbp]
 \begin{center}
 \begin{myenuma}
  \panel{\includegraphics[width=0.14\linewidth]{triangular_right.svg}}\label{fig:tri_right}\hp
  \panel{\includegraphics[width=0.14\linewidth]{triangular_left.svg}}\label{fig:tri_left}\hp
  \panel{\includegraphics[width=0.28\linewidth]{shortcut.svg}}\label{fig:shortcut}
 \end{myenuma}
 \end{center}
  \caption{Perturbations that increase the area.
  (\ref{fig:tri_right}) Perturbations that increase elements of $\M\pot$ above the diagonal and decrease the corresponding elements of $\M\dep$. It can no longer be used when $\M\dep$ is lower triangular, \ie depression must move synapses to ``more depressed'' states.
  (\ref{fig:tri_left}) Perturbations that decrease elements of $\M\pot$ below the diagonal and increase the corresponding elements of $\M\dep$. It can no longer be used when $\M\pot$ is upper triangular, \ie potentiation must move synapses to ``more potentiated'' states.
  (\ref{fig:shortcut}) Perturbation that decreases ``shortcut'' transitions and increases the bypassed ``direct'' transitions. It can no longer be used when there are only nearest-neighbor ``direct'' transitions.
  \label{fig:perts}}
\end{figure}

@@

\begin{figure}
 \begin{center}
 \parbox[t]{0.43\linewidth}{%
 \begin{myenuma}
   \panel{\includegraphics[width=0.85\linewidth]{env.svg}}\label{fig:envplot}
 \end{myenuma}}\hp
 \parbox[t]{0.44\linewidth}{%
 \begin{myenuma}
  \stepcounter{myenumai}
% \addtocounter{enumi}{1}
   \panel{\includegraphics[width=0.8\linewidth]{diffjump.svg}}\label{fig:diffjump}\\[0.1\linewidth]
   \panel{\includegraphics[width=0.8\linewidth]{multistate_sticky.svg}}\label{fig:sticky}
 \end{myenuma}}
 \end{center}
  \caption{The memory curve envelope for $N=100$, $M=12$.
  (\protect\protect\ref{fig:envplot}) An upper bound on the SNR at any time is shown in green.
  The red dashed curve shows the result of numerical optimization of synaptic models with random initialization.
  The solid red curve shows the highest SNR we have found with hand designed models.
  At early times these models are of the form shown in (\ref{fig:diffjump}) with different numbers of states, and all transition probabilities equal to $1$.
  At late times they are of the form shown in (\ref{fig:sticky}) with different values of $\varepsilon$.
  The model shown in (\ref{fig:sticky}) also saturates the area bound \cref{eq:max_area} in the limit $\varepsilon\to0$.
  \label{fig:env}}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\bibliographystyle{pnas-new}
%\bibliography{maths,neuro,markov}

%\end{article}
%\clearpage

\begin{strip}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\appendix
\begin{center}
  \textbf{\Large Supporting information}
\end{center}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In the \hyperref[sec:areacoeffincr]{\lcnamecref{sec:areacoeffincr} below}, we will show that the $c_k$ are non-decreasing, if we put the states in order of decreasing $\eta^+_k$.
This implies that the last term of the final expression in \cref{eq:shortcutarea} is non-negative.
If it is non-zero (there would need to be a lot of degeneracy for it to be zero), this would imply that $\mu\potdep _{g,g\pm m}>0$, which in turn implies that $\W\potdep _{g,g\pm m}=0$.
This would tell us that the process with the maximal area has to have a multi-state topology.

@@

In the \hyperref[sec:shortcuts]{previous \lcnamecref{sec:shortcuts}} we defined perturbations corresponding to Figure 3(c) of the main text.
In order to show that those perturbations increase the area, we must now show that the $c_k$ are non-decreasing, if we put the states in order of decreasing $\eta^+_k$.

@@

%\bibliographystyle{utcaps_sl}
\bibliography{maths,neuro,markov}
